---
title: "VMBGCC Poster Analysis & Figures"
subtitle: "Brain-Gut Coaching Class — ePoster for Academic Conference"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(scales)
library(patchwork)
library(ggalluvial)
library(effectsize)
library(boot)

opts_chunk$set(
  fig.width = 10, fig.height = 7, cache = FALSE,
  echo = FALSE, warning = FALSE, message = FALSE,
  results = "markup"
)
options(stringsAsFactors = FALSE)

baseDir <- "/Users/tedwards/Documents/projects/VM_brainGutCoaching"
setwd(baseDir)

dataOutputDir <- file.path(baseDir, "data/outputData")
plotDir <- file.path(baseDir, "figures")
dataDate <- "2026-03-19"
filenameSuffix <- paste0("VMBGCC.", dataDate)

source(file.path(baseDir, "code/VMBGCC_functions.R"))

bgcc.df <- readRDS(file.path(dataOutputDir, paste0(filenameSuffix, "_bgccClean.rds")))
thematic.df <- readRDS(file.path(dataOutputDir, paste0(filenameSuffix, "_thematicCoding.rds")))

cat("Loaded:", nrow(bgcc.df), "patients,", nrow(thematic.df), "thematic rows\n")

# ---- Poster colour palette and ggplot theme ----
posterColors <- list(
  pre = "#4477AA",
  post = "#EE6677",
  mainTheme = "#228833",
  subTheme = "#CCBB44",
  severity = c(
    "Remission" = "#228833", "Mild" = "#CCBB44",
    "Moderate" = "#EE6677", "Severe" = "#AA3377"
  ),
  themeGroup = c(
    "Positive Community Atmosphere" = "#4477AA",
    "Collaborative Engaging Educational Environment" = "#228833",
    "Equipping Patients with Hope and Actionable Tools" = "#EE6677"
  )
)

theme_poster <- function(base_size = 14) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      plot.title = element_text(size = base_size + 4, face = "bold", hjust = 0),
      plot.subtitle = element_text(size = base_size, hjust = 0, color = "grey30"),
      axis.title = element_text(size = base_size, face = "bold"),
      axis.text = element_text(size = base_size - 1),
      legend.title = element_text(size = base_size - 1, face = "bold"),
      legend.text = element_text(size = base_size - 2),
      panel.grid.minor = element_blank(),
      strip.text = element_text(size = base_size, face = "bold")
    )
}

# ---- Theme vectors and labels ----
# NOTE: Data column names retain legacy identifiers; display labels match the
# clinician-researcher consensus poster (overarching theme: Patient Empowerment).
mainThemes <- c("themePositiveSharedExperience", "themePatientEmpowerment",
  "themePatientActivation")
subThemes <- c("subIsolationReduced", "subValidation", "subSharedCommunity",
  "subChangeNegHealthcare", "subGratitude", "subEnjoymentPositive",
  "subContent", "subTeachingCoachingStyle", "subCoachingStructure",
  "subSelfEfficacy", "subAgencyKnowledgeSkills", "subAgencyActionableTools")
allThemes <- c(mainThemes, subThemes)

overarchingTheme <- "Patient Empowerment"

themeLabels <- c(
  themePositiveSharedExperience = "Positive Community Atmosphere",
  themePatientEmpowerment       = "Collaborative Engaging Educational Environment",
  themePatientActivation        = "Equipping Patients with Hope and Actionable Tools",
  subIsolationReduced       = "Isolation Reduced",
  subValidation             = "Validation",
  subSharedCommunity        = "Shared Community",
  subChangeNegHealthcare    = "Change from Negative Healthcare Experiences",
  subGratitude              = "Gratitude",
  subEnjoymentPositive      = "Positive Experience",
  subContent                = "Presentation Content",
  subTeachingCoachingStyle  = "Presenter Teaching/Coaching Style",
  subCoachingStructure      = "Virtual Coaching Class Format",
  subSelfEfficacy           = "Patient Self-Efficacy",
  subAgencyKnowledgeSkills  = "Patient Agency (Knowledge, Skills)",
  subAgencyActionableTools  = "Patient Agency (Actionable Tools)"
)

# Map sub-themes to their parent main theme for color grouping
subToMain <- c(
  subIsolationReduced       = "Positive Community Atmosphere",
  subValidation             = "Positive Community Atmosphere",
  subSharedCommunity        = "Positive Community Atmosphere",
  subChangeNegHealthcare    = "Positive Community Atmosphere",
  subGratitude              = "Positive Community Atmosphere",
  subEnjoymentPositive      = "Positive Community Atmosphere",
  subContent                = "Collaborative Engaging Educational Environment",
  subTeachingCoachingStyle  = "Collaborative Engaging Educational Environment",
  subCoachingStructure      = "Collaborative Engaging Educational Environment",
  subSelfEfficacy           = "Equipping Patients with Hope and Actionable Tools",
  subAgencyKnowledgeSkills  = "Equipping Patients with Hope and Actionable Tools",
  subAgencyActionableTools  = "Equipping Patients with Hope and Actionable Tools"
)
```

# Abstract Verification

```{r abstract-verify}
cat("=== Abstract Number Verification ===\n\n")
N <- nrow(bgcc.df)
cat("N =", N, "(abstract: 183)\n")
cat("Mean age =", round(mean(bgcc.df$age, na.rm = TRUE), 1),
  "(SD", round(sd(bgcc.df$age, na.rm = TRUE), 1), ") — abstract: 53.1 (SD 16.6)\n")
nF <- sum(bgcc.df$sex == "F", na.rm = TRUE)
cat("Female:", nF, "/", N, "=", round(100 * nF / N, 1), "% — abstract: 76.9%\n")
cat("Mean pre-IBS-SSS:", round(mean(bgcc.df$preIBSSSS, na.rm = TRUE), 1),
  "(SD", round(sd(bgcc.df$preIBSSSS, na.rm = TRUE), 1), ") — abstract: 228 (SD 119.5)\n")

nResp <- sum(bgcc.df$respondedToSurvey, na.rm = TRUE)
cat("Survey respondents:", nResp, "/", N, "=", round(100 * nResp / N, 1), "% — abstract: 109 (59.6%)\n")

cat("\nDate range:", as.character(min(bgcc.df$classDate, na.rm = TRUE)),
  "to", as.character(max(bgcc.df$classDate, na.rm = TRUE)),
  "— abstract: 6/1/2023–8/3/2025\n")

cat("\n** NOTE: Minor discrepancies likely reflect a data update after abstract submission. **\n")
cat("** Current pipeline values should be used for the poster. **\n")
```

# Phase A.0 — Selection Bias / Missingness Table

```{r selection-bias}
# Define cohort strata
bgcc.df$stratum_respondent <- bgcc.df$respondedToSurvey
bgcc.df$stratum_pairedIBS <- !is.na(bgcc.df$preIBSSSS) & !is.na(bgcc.df$postIBSSSS)
bgcc.df$stratum_pairedPHQ <- !is.na(bgcc.df$prePHQ2) & !is.na(bgcc.df$postPHQ2)
bgcc.df$stratum_pairedGAD <- !is.na(bgcc.df$preGAD7) & !is.na(bgcc.df$postGAD7)

# Merge thematic coding for intersection check
themeRespondents <- thematic.df %>%
  filter(!is.na(surveyLiked) & tolower(surveyLiked) != "did not respond")
respondentCodes <- themeRespondents$patientCode

bgcc.df$stratum_intersectionIBS <- bgcc.df$patientCode %in% respondentCodes &
  bgcc.df$stratum_pairedIBS

cat("=== Cohort Strata ===\n")
cat("A: All attendees         :", nrow(bgcc.df), "\n")
cat("B: Survey respondents    :", sum(bgcc.df$stratum_respondent), "\n")
cat("C: Paired IBS-SSS        :", sum(bgcc.df$stratum_pairedIBS), "\n")
cat("D: Paired PHQ-2          :", sum(bgcc.df$stratum_pairedPHQ), "\n")
cat("E: Paired GAD-7          :", sum(bgcc.df$stratum_pairedGAD), "\n")
cat("F: Respondent+IBS-SSS    :", sum(bgcc.df$stratum_intersectionIBS), "\n")
cat("G: Respondent+PHQ-2      :", sum(bgcc.df$patientCode %in% respondentCodes & bgcc.df$stratum_pairedPHQ), "\n")
cat("H: Respondent+GAD-7      :", sum(bgcc.df$patientCode %in% respondentCodes & bgcc.df$stratum_pairedGAD), "\n")

# Function to compare two groups on baseline characteristics
compareGroups <- function(df, groupVar, label) {
  g1 <- df[df[[groupVar]], ]
  g0 <- df[!df[[groupVar]], ]
  cat(sprintf("\n--- %s (n=%d) vs NOT (n=%d) ---\n", label, nrow(g1), nrow(g0)))

  # Age
  wAge <- wilcox.test(g1$age, g0$age)
  cat(sprintf("  Age: %.1f (%.1f) vs %.1f (%.1f), Wilcoxon p=%s\n",
    mean(g1$age, na.rm = TRUE), sd(g1$age, na.rm = TRUE),
    mean(g0$age, na.rm = TRUE), sd(g0$age, na.rm = TRUE),
    format.pval(wAge$p.value, digits = 3)))

  # Sex
  t1 <- table(g1$sex)
  t0 <- table(g0$sex)
  pctF1 <- 100 * sum(g1$sex == "F", na.rm = TRUE) / nrow(g1)
  pctF0 <- 100 * sum(g0$sex == "F", na.rm = TRUE) / nrow(g0)
  sexTab <- table(df$sex, df[[groupVar]])
  pSex <- tryCatch(chisq.test(sexTab)$p.value, error = function(e) fisher.test(sexTab)$p.value)
  cat(sprintf("  Female: %.1f%% vs %.1f%%, p=%s\n", pctF1, pctF0, format.pval(pSex, digits = 3)))

  # Baseline IBS-SSS
  wIBS <- tryCatch(wilcox.test(g1$preIBSSSS, g0$preIBSSSS)$p.value,
    error = function(e) NA)
  cat(sprintf("  Pre-IBS-SSS: %.1f (%.1f) vs %.1f (%.1f), p=%s\n",
    mean(g1$preIBSSSS, na.rm = TRUE), sd(g1$preIBSSSS, na.rm = TRUE),
    mean(g0$preIBSSSS, na.rm = TRUE), sd(g0$preIBSSSS, na.rm = TRUE),
    format.pval(wIBS, digits = 3)))

  # nDiagnoses
  wDx <- wilcox.test(g1$nDiagnoses, g0$nDiagnoses)
  cat(sprintf("  nDiagnoses: %.1f (%.1f) vs %.1f (%.1f), p=%s\n",
    mean(g1$nDiagnoses, na.rm = TRUE), sd(g1$nDiagnoses, na.rm = TRUE),
    mean(g0$nDiagnoses, na.rm = TRUE), sd(g0$nDiagnoses, na.rm = TRUE),
    format.pval(wDx$p.value, digits = 3)))

  # Baseline severity band (if available)
  sevTab1 <- table(g1$preIBSSSSBand)
  sevTab0 <- table(g0$preIBSSSSBand)
  if (sum(sevTab1) > 0 && sum(sevTab0) > 0) {
    cat("  IBS-SSS Severity (In group):", paste(names(sevTab1), sevTab1, collapse = ", "), "\n")
    cat("  IBS-SSS Severity (Not in group):", paste(names(sevTab0), sevTab0, collapse = ", "), "\n")
  }
}

compareGroups(bgcc.df, "stratum_respondent", "Survey Respondent")
compareGroups(bgcc.df, "stratum_pairedIBS", "Paired IBS-SSS")
compareGroups(bgcc.df, "stratum_intersectionIBS", "Respondent + Paired IBS-SSS")

# Clean up temporary columns
bgcc.df <- bgcc.df %>% select(-starts_with("stratum_"))
```

# Phase A.1-4 — Theme × Outcome Analysis

```{r theme-outcome-setup}
# Merge thematic coding with clinical data
themeOutcome.df <- bgcc.df %>%
  inner_join(themeRespondents %>% select(patientCode, all_of(allThemes)),
    by = "patientCode")

cat("Theme × outcome dataset:", nrow(themeOutcome.df), "patients\n")
cat("  With paired IBS-SSS:", sum(!is.na(themeOutcome.df$deltaIBSSSS)), "\n")
cat("  With paired PHQ-2:", sum(!is.na(themeOutcome.df$deltaPHQ2)), "\n")
cat("  With paired GAD-7:", sum(!is.na(themeOutcome.df$deltaGAD7)), "\n")
```

## IBS-SSS Responder Rate by Baseline Severity

```{r responder-by-severity}
ibsPaired <- bgcc.df %>%
  filter(!is.na(preIBSSSS) & !is.na(postIBSSSS)) %>%
  mutate(ibsResponder = deltaIBSSSS <= -50)

cat("=== IBS-SSS Responder Rate by Baseline Severity ===\n\n")
cat("Overall:", sum(ibsPaired$ibsResponder), "/", nrow(ibsPaired),
  "=", round(100 * mean(ibsPaired$ibsResponder), 1), "%\n\n")

for (band in c("Remission", "Mild", "Moderate", "Severe")) {
  sub <- ibsPaired %>% filter(preIBSSSSBand == band)
  if (nrow(sub) > 0) {
    nR <- sum(sub$ibsResponder)
    nT <- nrow(sub)
    pct <- round(100 * nR / nT, 1)
    ci <- if (nR > 0 && nR < nT) {
      ci <- prop.test(nR, nT, correct = FALSE)$conf.int
      sprintf("[%.1f%%, %.1f%%]", 100 * ci[1], 100 * ci[2])
    } else {
      "—"
    }
    cat(sprintf("  %-12s: %d/%d = %5.1f%% %s\n", band, nR, nT, pct, ci))
  }
}
cat("\nNote: Responder = ≥50-pt IBS-SSS decrease. Most meaningful for Moderate/Severe baseline.\n")
```

## Main Theme × Primary Outcome (Formal Tests)

```{r theme-ibs-formal}
# Formal tests: 3 main themes × deltaIBSSSS only
toIBS <- themeOutcome.df %>% filter(!is.na(deltaIBSSSS))
nIBS <- nrow(toIBS)
cat("=== Main Theme × deltaIBSSSS (N =", nIBS, ") ===\n\n")

if (nIBS >= 10) {
  mainThemeIBS <- data.frame(
    theme = character(), label = character(),
    n0 = integer(), n1 = integer(),
    medianDelta0 = numeric(), medianDelta1 = numeric(),
    meanDelta0 = numeric(), meanDelta1 = numeric(),
    wilcoxP = numeric(), rankBisR = numeric(),
    rCIlow = numeric(), rCIhigh = numeric(),
    stringsAsFactors = FALSE
  )

  for (theme in mainThemes) {
    g0 <- toIBS$deltaIBSSSS[toIBS[[theme]] == 0]
    g1 <- toIBS$deltaIBSSSS[toIBS[[theme]] == 1]
    if (length(g0) >= 3 && length(g1) >= 3) {
      wt <- wilcox.test(g1, g0, conf.int = TRUE)
      rb <- effectsize::rank_biserial(g1, g0)
      mainThemeIBS <- bind_rows(mainThemeIBS, data.frame(
        theme = theme, label = themeLabels[theme],
        n0 = length(g0), n1 = length(g1),
        medianDelta0 = median(g0), medianDelta1 = median(g1),
        meanDelta0 = mean(g0), meanDelta1 = mean(g1),
        wilcoxP = wt$p.value,
        rankBisR = rb$r_rank_biserial,
        rCIlow = rb$CI_low, rCIhigh = rb$CI_high,
        stringsAsFactors = FALSE
      ))
    }
  }
  mainThemeIBS$fdrP <- p.adjust(mainThemeIBS$wilcoxP, method = "BH")

  cat("Formal Mann-Whitney tests (FDR-corrected):\n")
  for (i in 1:nrow(mainThemeIBS)) {
    row <- mainThemeIBS[i, ]
    sig <- if (row$fdrP < 0.05) " *" else ""
    cat(sprintf("  %s: theme present (n=%d) median Δ=%.0f vs absent (n=%d) median Δ=%.0f\n",
      row$label, row$n1, row$medianDelta1, row$n0, row$medianDelta0))
    cat(sprintf("    r = %.3f [%.3f, %.3f], p = %.4f, FDR p = %.4f%s\n",
      row$rankBisR, row$rCIlow, row$rCIhigh, row$wilcoxP, row$fdrP, sig))
  }
}
```

## All Themes × All Outcomes (Effect Sizes — Descriptive)

```{r theme-outcome-effectsizes}
# Effect sizes for ALL 15 themes × 3 outcomes (no formal tests for sub-themes)
outcomes <- c("deltaIBSSSS", "deltaPHQ2", "deltaGAD7")
outcomeLabels <- c(deltaIBSSSS = "IBS-SSS Δ", deltaPHQ2 = "PHQ-2 Δ", deltaGAD7 = "GAD-7 Δ")

allEffects <- data.frame()

for (theme in allThemes) {
  for (outcome in outcomes) {
    sub <- themeOutcome.df %>% filter(!is.na(.data[[outcome]]))
    g0 <- sub[[outcome]][sub[[theme]] == 0]
    g1 <- sub[[outcome]][sub[[theme]] == 1]

    if (length(g0) >= 3 && length(g1) >= 3) {
      rb <- tryCatch(
        {
          res <- effectsize::rank_biserial(g1, g0)
          data.frame(r = res$r_rank_biserial, ciLow = res$CI_low, ciHigh = res$CI_high)
        },
        error = function(e) data.frame(r = NA, ciLow = NA, ciHigh = NA))

      allEffects <- bind_rows(allEffects, data.frame(
        theme = theme, themeLabel = themeLabels[theme],
        outcome = outcome, outcomeLabel = outcomeLabels[outcome],
        n0 = length(g0), n1 = length(g1),
        meanDelta0 = mean(g0), meanDelta1 = mean(g1),
        medDelta0 = median(g0), medDelta1 = median(g1),
        r = rb$r, rCIlow = rb$ciLow, rCIhigh = rb$ciHigh,
        type = ifelse(theme %in% mainThemes, "Main Theme", "Sub-Theme"),
        parentTheme = ifelse(theme %in% mainThemes, themeLabels[theme], subToMain[theme]),
        stringsAsFactors = FALSE
      ))
    }
  }
}

cat("Computed", nrow(allEffects), "effect sizes (", length(unique(allEffects$theme)),
  "themes ×", length(outcomes), "outcomes)\n\n")

# Print summary table
cat("=== Effect Size Summary (rank-biserial r) ===\n")
cat("Negative r = theme-present group has MORE negative delta (more improvement)\n\n")
for (out in outcomes) {
  cat(sprintf("--- %s ---\n", outcomeLabels[out]))
  sub <- allEffects %>% filter(outcome == out) %>% arrange(r)
  for (i in 1:nrow(sub)) {
    row <- sub[i, ]
    marker <- ifelse(row$type == "Main Theme", " [MAIN]", "")
    cat(sprintf("  %-35s: r = %+.3f [%+.3f, %+.3f] (n=%d/%d)%s\n",
      row$themeLabel, row$r, row$rCIlow, row$rCIhigh, row$n1, row$n0, marker))
  }
  cat("\n")
}
```

## Effect Size Heatmap (Poster Panel E)

```{r effect-heatmap, fig.width=12, fig.height=8}
# Order themes: main themes first, then sub-themes grouped by parent
themeOrder <- c(
  # Main themes
  "Positive Shared Experience", "Patient Empowerment", "Patient Activation",
  # Sub-themes under Positive Shared Experience
  "Isolation Reduced", "Validation", "Shared Community",
  "Change from Neg. Healthcare", "Gratitude", "Enjoyment / Positive",
  # Sub-themes under Patient Empowerment
  "Content", "Teaching / Coaching Style", "Coaching Structure",
  # Sub-themes under Patient Activation
  "Self-Efficacy", "Agency (Knowledge/Skills)", "Agency (Actionable Tools)"
)

heatData <- allEffects %>%
  mutate(
    themeLabel = factor(themeLabel, levels = rev(themeOrder)),
    outcomeLabel = factor(outcomeLabel, levels = c("IBS-SSS Δ", "PHQ-2 Δ", "GAD-7 Δ")),
    sigMarker = ifelse(!is.na(rCIlow) & !is.na(rCIhigh) &
      ((rCIlow > 0 & rCIhigh > 0) | (rCIlow < 0 & rCIhigh < 0)), "*", "")
  )

figHeatmap <- ggplot(heatData, aes(x = outcomeLabel, y = themeLabel, fill = r)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%+.2f%s", r, sigMarker)),
    size = 4, color = "black") +
  scale_fill_gradient2(
    low = "#228833", mid = "white", high = "#AA3377",
    midpoint = 0, limits = c(-1, 1),
    name = "Rank-\nbiserial r"
  ) +
  labs(
    x = NULL, y = NULL,
    title = "Exploratory: Theme Presence × Clinical Outcome Change",
    subtitle = "Negative r = theme-present group improved more | * = 95% CI excludes zero"
  ) +
  theme_poster(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid = element_blank()
  )

print(figHeatmap)
savePlot(figHeatmap, plotDir, "poster_themeOutcomeHeatmap", height = 8, width = 10)
```

## Logistic Regression: IBS-SSS Responder ~ Themes

```{r responder-regression}
# Intersection: respondents + paired IBS-SSS
toModel <- themeOutcome.df %>%
  filter(!is.na(deltaIBSSSS)) %>%
  mutate(ibsResponder = as.integer(deltaIBSSSS <= -50))

nModel <- nrow(toModel)
nResp <- sum(toModel$ibsResponder)
cat("=== Logistic Regression: Responder ~ Themes (N =", nModel, ", responders =", nResp, ") ===\n\n")

if (nModel >= 20 && nResp >= 5 && (nModel - nResp) >= 5) {
  # Check if we have enough events per predictor (rule of ~10 EPV)
  nPredictors <- 3  # main themes only
  epv <- min(nResp, nModel - nResp) / nPredictors
  cat("Events per variable:", round(epv, 1), "\n")

  if (epv >= 5) {
    # Adjusted model: themes + baseline + demographics
    nPredFull <- 6
    epvFull <- min(nResp, nModel - nResp) / nPredFull
    cat("Full model EPV:", round(epvFull, 1), "\n")

    if (epvFull >= 5) {
      glmFull <- glm(ibsResponder ~ themePositiveSharedExperience +
        themePatientEmpowerment + themePatientActivation +
        preIBSSSS + age + sex,
      data = toModel, family = binomial)
      cat("\nFull model (themes + covariates):\n")
      print(summary(glmFull)$coefficients)

      # Odds ratios with CIs
      orCI <- exp(cbind(OR = coef(glmFull), confint.default(glmFull)))
      cat("\nOdds Ratios:\n")
      print(round(orCI, 3))
    } else {
      cat("EPV too low for full model — fitting unadjusted model\n")
    }

    # Unadjusted models (one theme at a time)
    cat("\n--- Unadjusted ORs (one theme at a time) ---\n")
    for (theme in mainThemes) {
      glmUnadj <- glm(as.formula(paste("ibsResponder ~", theme)),
        data = toModel, family = binomial)
      orCI <- exp(cbind(OR = coef(glmUnadj), confint.default(glmUnadj)))
      pval <- summary(glmUnadj)$coefficients[2, 4]
      cat(sprintf("  %s: OR = %.2f [%.2f, %.2f], p = %.3f\n",
        themeLabels[theme], orCI[2, 1], orCI[2, 2], orCI[2, 3], pval))
    }
  } else {
    cat("Too few events per variable for logistic regression.\n")
    cat("Reporting Fisher exact tests instead.\n\n")
    for (theme in mainThemes) {
      tab <- table(toModel[[theme]], toModel$ibsResponder)
      ft <- fisher.test(tab)
      cat(sprintf("  %s: OR = %.2f [%.2f, %.2f], p = %.3f\n",
        themeLabels[theme], ft$estimate, ft$conf.int[1], ft$conf.int[2], ft$p.value))
    }
  }
} else {
  cat("Insufficient sample size for regression (N =", nModel, ", responders =", nResp, ").\n")
  cat("Descriptive comparison only.\n")
}
```

## Linear Model: deltaIBSSSS ~ Themes

```{r delta-lm}
if (nModel >= 20) {
  cat("=== Linear Model: deltaIBSSSS ~ themes + covariates ===\n\n")

  lmFull <- lm(deltaIBSSSS ~ themePositiveSharedExperience +
    themePatientEmpowerment + themePatientActivation +
    preIBSSSS + age + sex,
  data = toModel)

  cat("Model summary:\n")
  print(summary(lmFull))

  # Standardized coefficients
  cat("\nCoefficients with 95% CIs:\n")
  coefCI <- cbind(
    Estimate = coef(lmFull),
    confint(lmFull)
  )
  print(round(coefCI, 2))
}
```

# Phase A.5 — Micro-Goal Descriptive

```{r microgoals, fig.width=10, fig.height=6}
goalsText <- bgcc.df$microGoals[!is.na(bgcc.df$microGoals)]
nGoals <- length(goalsText)
cat("Patients with micro-goals:", nGoals, "\n\n")

goalCategories <- list(
  "Exercise / Movement" = c("exercise", "walk", "walking", "yoga", "stretch",
    "move", "movement", "physical", "hike", "run", "swim", "gym", "steps"),
  "Mindfulness / Meditation" = c("mindful", "meditat", "breath", "relax",
    "calm", "deep breath", "diaphragm"),
  "Diet / Nutrition" = c("diet", "eat", "food", "meal", "fiber", "water",
    "hydrat", "nutrition", "fodmap", "drink"),
  "Sleep / Rest" = c("sleep", "rest", "bedtime", "nap"),
  "Social Activity" = c("social", "friend", "family", "connect", "community",
    "group", "talk", "spend time"),
  "Screen Time Reduction" = c("screen", "phone", "device", "tv",
    "social media", "limit"),
  "Stress Management" = c("stress", "journal", "writing", "gratitude",
    "self-care", "selfcare"),
  "App / Tool Use" = c("nerva", "mahana", "calmigo", "app")
)

goalCounts <- sapply(goalCategories, function(keywords) {
  sum(sapply(goalsText, function(txt) any(str_detect(tolower(txt), keywords))))
})

goalData <- data.frame(
  category = names(goalCounts),
  n = as.integer(goalCounts),
  pct = 100 * goalCounts / nGoals,
  stringsAsFactors = FALSE
) %>% arrange(desc(n))
goalData$category <- factor(goalData$category, levels = rev(goalData$category))

cat("Micro-goal categories:\n")
for (i in 1:nrow(goalData)) {
  cat(sprintf("  %-25s: %3d (%.1f%%)\n",
    as.character(goalData$category[i]), goalData$n[i], goalData$pct[i]))
}

figGoals <- ggplot(goalData, aes(x = category, y = pct)) +
  geom_col(fill = posterColors$mainTheme, width = 0.7) +
  geom_text(aes(label = sprintf("%d (%.0f%%)", n, pct)),
    hjust = -0.1, size = 4) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.3))) +
  labs(x = NULL, y = "% of patients with micro-goals",
    title = sprintf("Patient-Set Micro-Goals (N=%d)", nGoals),
    subtitle = "Keyword-based categorization of self-directed goals set during class") +
  theme_poster()

print(figGoals)
savePlot(figGoals, plotDir, "poster_microGoals", height = 6, width = 10)
```

# Phase B — Poster Figures

## Panel B: Gardner-Altman Estimation Plot (IBS-SSS)

```{r gardner-altman, fig.width=10, fig.height=7}
ibsPaired <- bgcc.df %>%
  filter(!is.na(preIBSSSS) & !is.na(postIBSSSS)) %>%
  mutate(
    ibsResponder = deltaIBSSSS <= -50,
    respLabel = ifelse(ibsResponder, "Responder (≥50-pt decrease)", "Non-responder")
  )

nIBS <- nrow(ibsPaired)
meanDelta <- mean(ibsPaired$deltaIBSSSS)
medianDelta <- median(ibsPaired$deltaIBSSSS)

# Bootstrap 95% CI for mean delta
set.seed(42)
bootMean <- boot(ibsPaired$deltaIBSSSS, function(d, i) mean(d[i]), R = 2000)
bootCI <- boot.ci(bootMean, type = "perc")$percent[4:5]

# Paired data in long format
ibsLong <- ibsPaired %>%
  select(patientCode, preIBSSSS, postIBSSSS, ibsResponder) %>%
  pivot_longer(cols = c(preIBSSSS, postIBSSSS),
    names_to = "timepoint", values_to = "score") %>%
  mutate(
    time = factor(ifelse(grepl("pre", timepoint), "Pre-class", "Post-class"),
      levels = c("Pre-class", "Post-class"))
  )

# Left panel: paired lines
pLeft <- ggplot(ibsLong, aes(x = time, y = score)) +
  geom_line(aes(group = patientCode, color = ibsResponder),
    alpha = 0.35, linewidth = 0.4) +
  geom_point(aes(color = ibsResponder), alpha = 0.3, size = 1) +
  stat_summary(aes(group = 1), fun = mean, geom = "line",
    color = "black", linewidth = 2) +
  stat_summary(aes(group = 1), fun = mean, geom = "point",
    color = "black", size = 4, shape = 18) +
  geom_hline(yintercept = c(75, 175, 300), linetype = "dashed",
    color = c("#228833", "#CCBB44", "#AA3377"), alpha = 0.5) +
  scale_color_manual(values = c("TRUE" = "#228833", "FALSE" = "#AA3377"),
    labels = c("TRUE" = "Responder", "FALSE" = "Non-responder"),
    name = NULL) +
  labs(x = NULL, y = "IBS-SSS Score") +
  theme_poster() +
  theme(legend.position = "bottom")

# Right panel: delta distribution with mean + CI
pRight <- ggplot(ibsPaired, aes(x = 1, y = deltaIBSSSS)) +
  geom_hline(yintercept = 0, linetype = "solid", color = "grey50") +
  geom_hline(yintercept = -50, linetype = "dotted", color = "#228833", linewidth = 0.8) +
  geom_jitter(aes(color = ibsResponder), width = 0.15, alpha = 0.4, size = 2) +
  # Mean + CI
  annotate("rect", xmin = 0.6, xmax = 1.4,
    ymin = bootCI[1], ymax = bootCI[2],
    fill = "black", alpha = 0.15) +
  annotate("segment", x = 0.6, xend = 1.4,
    y = meanDelta, yend = meanDelta,
    color = "black", linewidth = 1.5) +
  annotate("text", x = 1.55, y = meanDelta,
    label = sprintf("Mean: %.0f\n[%.0f, %.0f]", meanDelta, bootCI[1], bootCI[2]),
    hjust = 0, size = 4, fontface = "bold") +
  scale_color_manual(values = c("TRUE" = "#228833", "FALSE" = "#AA3377"), guide = "none") +
  scale_x_continuous(limits = c(0.3, 2.2), breaks = NULL) +
  labs(x = NULL, y = "IBS-SSS Change\n(Post - Pre)") +
  theme_poster() +
  theme(axis.text.x = element_blank())

figEstimation <- pLeft + pRight +
  plot_layout(widths = c(2, 1)) +
  plot_annotation(
    title = sprintf("IBS-SSS Pre/Post Change (N=%d paired)", nIBS),
    subtitle = sprintf("Responders (≥50-pt decrease): %d/%d (%.0f%%) | Dotted line = clinically meaningful threshold",
      sum(ibsPaired$ibsResponder), nIBS, 100 * mean(ibsPaired$ibsResponder)),
    theme = theme_poster()
  )

print(figEstimation)
savePlot(figEstimation, plotDir, "poster_ibsEstimation", height = 7, width = 12)
```

## Panel C: Alluvial with Responder Annotations

```{r alluvial-annotated, fig.width=10, fig.height=7}
alluvialData <- ibsPaired %>%
  count(preIBSSSSBand, postIBSSSSBand) %>%
  rename(Pre = preIBSSSSBand, Post = postIBSSSSBand, Freq = n)

# Responder rate by baseline severity
respBySev <- ibsPaired %>%
  group_by(preIBSSSSBand) %>%
  summarise(
    n = n(), nResp = sum(ibsResponder),
    pctResp = round(100 * nResp / n, 0),
    .groups = "drop"
  )

figAlluvial <- ggplot(alluvialData,
  aes(axis1 = Pre, axis2 = Post, y = Freq)) +
  geom_alluvium(aes(fill = Pre), width = 1 / 3, alpha = 0.7) +
  geom_stratum(width = 1 / 3, fill = "grey90", color = "grey50") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 4.5) +
  scale_x_discrete(limits = c("Pre-class", "Post-class"),
    expand = c(0.15, 0.05)) +
  scale_fill_manual(values = posterColors$severity, name = "Pre-class\nSeverity") +
  labs(
    y = "Number of Patients",
    title = sprintf("IBS-SSS Severity Transitions (N=%d)", nIBS),
    subtitle = paste0(
      paste(sprintf("%s %d%% (%d/%d)", respBySev$preIBSSSSBand,
        respBySev$pctResp, respBySev$nResp, respBySev$n)[1:2], collapse = " | "),
      "\n",
      paste(sprintf("%s %d%% (%d/%d)", respBySev$preIBSSSSBand,
        respBySev$pctResp, respBySev$nResp, respBySev$n)[3:4], collapse = " | "))
  ) +
  theme_poster() +
  theme(legend.position = "right",
    plot.subtitle = element_text(size = 12))

print(figAlluvial)
savePlot(figAlluvial, plotDir, "poster_ibsAlluvial", height = 7, width = 10)
```

## Panel D: Theme Prevalence (Poster-Formatted)

```{r theme-prevalence-poster, fig.width=12, fig.height=8}
nThemeResp <- nrow(themeRespondents)

themeData <- data.frame(
  theme = allThemes,
  label = themeLabels[allThemes],
  n = sapply(allThemes, function(t) sum(themeRespondents[[t]] == 1)),
  stringsAsFactors = FALSE
) %>%
  mutate(
    pct = 100 * n / nThemeResp,
    type = ifelse(theme %in% mainThemes, "Main Theme", "Sub-Theme"),
    parentTheme = ifelse(theme %in% mainThemes, themeLabels[theme], subToMain[theme])
  )

# Wilson CIs
for (i in 1:nrow(themeData)) {
  ci <- prop.test(themeData$n[i], nThemeResp, correct = FALSE)$conf.int
  themeData$ciLow[i] <- 100 * ci[1]
  themeData$ciHigh[i] <- 100 * ci[2]
}

# Order: main themes first at top, then sub-themes grouped by parent, sorted by prevalence
themeData <- themeData %>%
  mutate(
    label = factor(label, levels = rev(themeOrder))
  )

figThemePrevalence <- ggplot(themeData, aes(x = label, y = pct, fill = parentTheme)) +
  geom_col(width = 0.7, alpha = 0.85) +
  geom_errorbar(aes(ymin = ciLow, ymax = ciHigh), width = 0.2, linewidth = 0.5) +
  geom_text(aes(label = sprintf("%d (%.0f%%)", n, pct)),
    hjust = -0.1, size = 3.8) +
  coord_flip() +
  scale_fill_manual(values = posterColors$themeGroup, name = "Theme Group") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.25))) +
  labs(
    x = NULL, y = "Prevalence (% of respondents)",
    title = sprintf("Thematic Analysis Results (N=%d respondents)", nThemeResp),
    subtitle = "Binary per participant (≥1 mention = present) | Error bars = Wilson 95% CI"
  ) +
  theme_poster() +
  theme(legend.position = "bottom")

print(figThemePrevalence)
savePlot(figThemePrevalence, plotDir, "poster_themePrevalence", height = 8, width = 12)
```

## PHQ-2 and GAD-7 Pre/Post (Supporting)

```{r phq-gad-prepost, fig.width=12, fig.height=6}
makeEstimationPlot <- function(df, preCol, postCol, scoreName, thresholds = NULL,
                               threshColors = NULL) {
  paired <- df %>%
    filter(!is.na(!!sym(preCol)) & !is.na(!!sym(postCol))) %>%
    mutate(delta = !!sym(postCol) - !!sym(preCol))

  nPaired <- nrow(paired)
  meanD <- mean(paired$delta)

  set.seed(42)
  bootM <- boot(paired$delta, function(d, i) mean(d[i]), R = 2000)
  bCI <- boot.ci(bootM, type = "perc")$percent[4:5]

  long <- paired %>%
    select(patientCode, !!sym(preCol), !!sym(postCol)) %>%
    pivot_longer(cols = c(!!sym(preCol), !!sym(postCol)),
      names_to = "tp", values_to = "score") %>%
    mutate(time = factor(ifelse(grepl("pre", tp), "Pre", "Post"), levels = c("Pre", "Post")))

  p <- ggplot(long, aes(x = time, y = score)) +
    geom_line(aes(group = patientCode), alpha = 0.15, color = "grey50") +
    geom_point(alpha = 0.15, size = 0.8, color = "grey50") +
    stat_summary(aes(group = 1), fun = mean, geom = "line",
      color = "black", linewidth = 1.5) +
    stat_summary(aes(group = 1), fun = mean, geom = "point",
      color = "black", size = 3, shape = 18) +
    labs(
      x = NULL, y = scoreName,
      title = sprintf("%s (N=%d)", scoreName, nPaired),
      subtitle = sprintf("Mean Δ = %.2f [%.2f, %.2f]", meanD, bCI[1], bCI[2])
    ) +
    theme_poster(base_size = 12)

  if (!is.null(thresholds)) {
    for (i in seq_along(thresholds)) {
      p <- p + geom_hline(yintercept = thresholds[i], linetype = "dashed",
        color = threshColors[i], alpha = 0.5)
    }
  }
  p
}

pPHQ <- makeEstimationPlot(bgcc.df, "prePHQ2", "postPHQ2", "PHQ-2",
  thresholds = 3, threshColors = "#AA3377")
pGAD <- makeEstimationPlot(bgcc.df, "preGAD7", "postGAD7", "GAD-7",
  thresholds = c(5, 10, 15), threshColors = c("#CCBB44", "#EE6677", "#AA3377"))

figPHQGAD <- pPHQ + pGAD +
  plot_annotation(
    title = "Supporting Outcomes: PHQ-2 and GAD-7",
    theme = theme_poster()
  )

print(figPHQGAD)
savePlot(figPHQGAD, plotDir, "poster_phqGadPrePost", height = 6, width = 12)
```

# Summary of Key Findings

```{r summary}
cat("============================================================\n")
cat("POSTER KEY FINDINGS SUMMARY\n")
cat("============================================================\n\n")

cat("COHORT: N =", nrow(bgcc.df), "patients,", nrow(themeRespondents), "survey respondents\n")
cat("  Mean age:", round(mean(bgcc.df$age, na.rm = TRUE), 1),
  "| Female:", round(100 * sum(bgcc.df$sex == "F", na.rm = TRUE) / nrow(bgcc.df), 1), "%\n\n")

cat("IBS-SSS (N =", nrow(ibsPaired), "paired):\n")
cat("  Pre: mean", round(mean(ibsPaired$preIBSSSS), 1),
  "→ Post: mean", round(mean(ibsPaired$postIBSSSS), 1), "\n")
cat("  Mean change:", round(meanDelta, 1), "[", round(bootCI[1], 1), ",", round(bootCI[2], 1), "]\n")
cat("  Responders:", sum(ibsPaired$ibsResponder), "/", nIBS,
  "(", round(100 * mean(ibsPaired$ibsResponder), 1), "%)\n\n")

cat("THEMATIC (N =", nThemeResp, "respondents):\n")
for (t in mainThemes) {
  n <- sum(themeRespondents[[t]] == 1)
  cat(sprintf("  %-35s: %d (%.1f%%)\n", themeLabels[t], n, 100 * n / nThemeResp))
}

cat("\nTHEME × OUTCOME INTERSECTION SIZES:\n")
cat("  Respondent + paired IBS-SSS: N =", nrow(toIBS), "\n")
cat("  Respondent + paired PHQ-2:   N =", sum(!is.na(themeOutcome.df$deltaPHQ2)), "\n")
cat("  Respondent + paired GAD-7:   N =", sum(!is.na(themeOutcome.df$deltaGAD7)), "\n")

cat("\nABSTRACT DISCREPANCIES (current data vs abstract):\n")
cat("  N: 182 vs 183 (1 patient difference)\n")
cat("  Female %: 76.4% vs 76.9%\n")
cat("  Respondents: 112 vs 109\n")
cat("  Theme %s are close but not exact (likely reflects updated consensus coding)\n")
cat("  Date range: 2023-12-15 to 2025-07-11 vs abstract 6/1/2023–8/3/2025\n")
cat("  IBS-C 41.8% vs 29.0% — LARGE DISCREPANCY (may be denominator difference)\n")
cat("  Functional abd pain 7.1% vs 16.4% — LARGE DISCREPANCY\n")
```

```{r cleanup}
# Remove temporary verification script
if (file.exists(file.path(baseDir, "code/abstract_verify.R"))) {
  file.remove(file.path(baseDir, "code/abstract_verify.R"))
}
```
